Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-01-18 2_SSL_EES/article/article_main.tex
updated figure description
added content to main document

\begin{figure*}[h]
 \centering
 \includegraphics[height=7cm]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history from 1850 to 2020. Data shown for best performers by year of market introduction, with dashed lines. A multi-order polynomial fit was used to indicate the average rate of change as dashed lines. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_max^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID – High-Intensity Discharge; CFL – Compact Fluorescent Lamp; Hal. – Halogen, Incd. – Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, section , for the full list of data sources. }
 \label{fgr:history_efficacy}
\end{figure*}

@@

Several descriptive papers include very useful information on some of the scientific breakthroughs that have contributed to LED progress \cite{krames2007status}, \cite{Phillips2007}, \cite{Bierhuizen2007}, \cite{Nakamura2013}, \cite{feezell2018invention} and \cite{Taki2019}. Additional  studies providing more detail on specific advances in LED technology were also published by authors working for key industry actors Lumileds\cite{MuellerMach2005}, \cite{Shchekin2006}, \cite{lumi2015lumi}, \cite{Bhardwaj2017} and Osram (now amsOsram) \cite{Haerle2004}, \cite{Baur2009}, \cite{laubsch2009high}, \cite{hahn2014development}. However, these studies typically include limited information on the macro chip-design and cover disparate aspects of the technology over different time periods. Another important source of historical information comes from patents, which cover the entirety of the device architecture or manufacturing process, including those by Lumileds \cite{margalith2011thin}, Samsung \cite{jung2014phos} \cite{cha2019semiconductor} and Soraa \cite{cich2017high}.

@@

>>>>

\begin{figure*}[h]
 \centering
 \includegraphics[height=7cm]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history from 1850 to 2020. Data shown for best performers by year of market introduction, with dashed lines. Dashed lines indicate the average improvement calculated from a 3rd-order polynomial fit to the data. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_max^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID – High-Intensity Discharge; CFL – Compact Fluorescent Lamp; Hal. – Halogen, Incd. – Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, section , for the full list of data sources. }
 \label{fgr:history_efficacy}
\end{figure*}

@@

Several descriptive publications provide a very useful overview of some of the scientific breakthroughs  that have contributed to LED progress \cite{krames2007status}\cite{Phillips2007}\cite{Bierhuizen2007}\cite{Nakamura2013}\cite{feezell2018invention}\cite{Taki2019} . Additional  studies providing more detail the origin and impact of specific advances in LED technology as well as their integration into device manufacturing were published by authors working for key industry actors Lumileds\cite{MuellerMach2005}\cite{Shchekin2006}\cite{lumi2015lumi}\cite{Bhardwaj2017} and Osram (now amsOsram) \cite{Haerle2004}\cite{Baur2009}\cite{laubsch2009high}\cite{hahn2014development}. However, these studies typically include limited information on the macro-level chip-design and cover disparate aspects of the technology over different time periods. Another important source of historical information comes from patents, which cover the entirety of the device architecture or manufacturing process, including those by Lumileds \cite{margalith2011thin}, Samsung \cite{jung2014phos} \cite{cha2019semiconductor} and Soraa \cite{cich2017high}.

@@

\section{Metrics for technological progress in LEDs}

Investigating the sources of rapid progress in LED technology over the past decades, in which white LEDs have come to dominate the lighting market \cite{zissis2021}, requires selecting appropriate metrics for tracking and quantifying this progress. The choice of metrics affects both what data sources can be used in the analysis and which research methodologies can be used to calculate and analyse such metrics. We select the metrics based on the following two general criteria. First, we focus only on metrics widely accepted and reported in industry because metrics proposed in the scientific literature but not reported by device manufacturers cannot be used to compare the performance of commercial LED devices over time. Second, the chosen metrics must be useful for understanding the impact of individual technological improvements on relevant performance and cost characteristics.

Historically, the progress in LED technology is commonly described by pointing to impressive improvements in LED device performance (i.e., brightness and electrical efficiency) and manufacturing cost reductions \cite{Taki2019}. However, solid-state lighting performance is characterized not only by LED brightness or electrical efficiency. Consumer experience, including the perceived temperature of a white light source and its ability to faithfully render colours, has also played a significant part in the market adoption of new lighting technologies \cite{Menanteau2000}\cite{Sandahl2006}\cite{CAIRD2008}\cite{murphy2012governing} and received substantial attention in LED research and development efforts \cite{azevedo2009transition}\cite{cho2017white}. Therefore, a comprehensive analysis of the evolution in white LED technology must take into consideration advances in: 1) physical device performance; 2) consumer experience; and 3) LED device manufacturing costs. Next, we introduce and discuss the metrics that we use to track progress in each of these three areas.

\subsection{Device Performance Metrics}

Initially, as LED devices developed, the primary metric of progress in solid-state lighting was typically the luminous flux, as reflected in the so-called “Haitz’s Law” \cite{haitz1999case}\cite{haitz2011solid}. The luminous flux (total brightness) of early white LEDs was too small to allow for the economical combination of multiple LEDs into lamps for general illumination purposes. In 2000, highest performing devices yielded around 10lm, just below the output of a candle as defined in the unit candela (1cd=12.57lm) \cite{haitz2011solid}. Today, however, highest performing devices yield in excess of 1600lm, the equivalent of a 100W incandescent bulb \cite{cree2020bright}. LED brightness has become sufficient to enable the construction of lamps from multiple LED devices, with design improvements focusing on higher efficiency and quality of light instead of brightness. Even though some scientific publications and industry periodicals continue to focus on brightness as a metric of progress in lighting, we find that, by now, it fails to capture the complexity of the multitude of efficiency improvements that have been driving overall LED efficiency \cite{weinold2021compound}.

\subsection{Consumer Experience Metrics}

The perceived quality of light is entirely determined by the emission spectrum of a light source \cite{ies_handbook}. Any metric relevant to customer experience can thus be calculated from the spectrum alone. The spectrum of an LED light source is determined by the emission wavelength of the LED itself and the absorption and emission spectra  of the downconversion phosphor used in the device. It is typically included in the product datasheets provided by manufacturers, which enables the calculation of all relevant spectrum-based metrics for these devices. Based on the prevalence in scholarly literature and industry publications, for this study we choose two consumer experience metrics: Colour Rendering Index (CRI) and Colour Temperature. We do not consider flicker, the unintended high-frequency temporal modulation of light, which is another important consumer experience metric for lighting and a subject of recent regulation by the European Union /cite{weinold2020long}. This effect is caused not by LEDs themselves, but rather by inadequately designed electrical ballasts \cite{Lehman2014}. It is thus beyond the scope of this work.

The Colour Rendering Index (CRI) of a light source describes its ability to render the colours of an object faithfully when compared to illumination under a reference light source, such as standard daylight \cite{khan2015led}. The way it is calculated is defined by the International Illumination Commission (CIE) \cite{cie_cri_1995}. CRI has certain limitations when applied to solid-state light sources \cite{david2013cri}. However, despite repeated attempts at constructing more elaborate colour rendering metrics \cite{Houser2013}, CRI has remained the de-facto industry standard for describing colour fidelity of light sources \cite{DoE2016LED}. High colour rendering performance of lighting is a requirement in workplace environments, retail stores, clinical operating environments and art exhibitions \cite{khanh2017color}. It should be noted that some niche applications prioritize high colour saturation over high CRI, for instance, in food display or fabric retail applications \cite{david2013cri}. However, these niche applications remain outside our focus on general illumination. Due to a broad availability and importance of CRI data for consumers of various LED lighting sources, we adopt CRI as the key metric to track progress in consumer experience in LED lighting, despite its limitations.

The Colour Temperature of a light source describes the equivalent temperature of an ideal black body which emits light of a colour comparable to that of the light source \cite{commission2011cie}. Warm white light sources are widely used in general illumination, while cold white light sources are used in workplace illumination and outdoor lighting. Early white LEDs produced only cool white light \cite{mueller2000light}. The introduction of first commercial warm white LED light bulbs played a significant part in increasing adoption of LED-based lighting among consumers, as their spectrum more closely resembled the warm white colour temperature of incandescent light bulbs \cite{al2016optics}. For this reason, we also adopt colour temperature as a metric for tracking progress in consumer experience in LED lighting.

\subsection{Manufacturing Cost}

In selecting metrics for tracking the progress in manufacturing cost reductions in LED, we must first highlight  the complexity of this task. Access to manufacturing cost data at the chip level is usually restricted due to its proprietary nature and is available only for selected products. Using sales price information instead of the cost for the same purpose could be in some cases a promising alternative, as prices can be easily obtained for recent LED products directly from manufacturers. However, historical data on prices for LEDs is as difficult to obtain as for costs. In addition, sales price includes components not relevant to the progress in technology, such as profit margins and overhead costs, and is affected by policies such as rebates and purchase subsidies. Historical information on these factors and price components, as well as their impact on the manufacturing cost for each LED product under consideration is even harder to obtain, making the use of LED sales prices as a direct metric of technology progress very difficult in practice.

\section{Methods and Data Collection}

The evolution of LED device architecture and performance as well as the progress in understanding the underlying physical phenomena are well covered in the scholarly literature and patents, as described in section 2. However, information provided in such sources is insufficient for our goals on at least three accounts. First, existing work focusses  only on selected performance parameters or overall device efficiency, rather than on providing a comprehensive coverage of the whole device sub-efficiencies for a particular LED product or design. Scientific publications also do not always disclose the underlying device architecture or the features responsible for the gains in performance. Second, not all relevant innovations are patented \cite{Pakes_1980}\cite{ Fontana_2013}. In the case of LED patents in particular, our interviews with industry experts suggest that the propensity to patent is the highest for knowledge related to macroscopic device architecture and chemical composition of phosphors, and the lowest for knowledge related to manufacturing process improvements and microscopic chip architecture that is difficult to reconstruct by reverse engineering. This means that relying only on patent literature would bias results by unduly emphasizing some focus areas and de-emphasizing others. Third, scientific publications and patents typically focus on experimental devices, rather than commercial products. While new LED features, designs and manufacturing methods reported in these sources can potentially result in significant performance gains or cost reductions, it is difficult to ascertain if these improvements have since been adopted in industry.  Furthermore, information on LED manufacturing cost and the effect of process improvements on the total cost is highly proprietary. Estimates are occasionally reported in the scientific literature and company publications, but these often do not disclose which parts of the manufacturing process are responsible for the largest contribution to the overall cost, or which improvements led to cost reductions.

To overcome the limitations of these different methods for understanding technological progress, in this study we rely on a multi-method approach to data collection and analysis that combines information obtained from a review of the primary scientific literature and relevant patents with our own computations of device sub-efficiencies, information gained from industry publications, semi-structured interviews with experts from academia and industry, and bottom-up manufacturing cost modelling. We provide the details of implementation of each of these methods below.

\subsection{Systematic Literature Review}

We collected data on LED performance and characteristics in a systematic literature review that included scientific publications, patents, conference proceedings from the largest semiconductor and optoelectronics conferences, industry periodicals and roadmaps, as well as company presentations and reports. This review was structured around the three main goals: 1) tracking the evolution of LED technology over time as indicated by three groups of progress metrics selected above in the “Metrics” section; 2) identifying individual innovations that contributed to this evolution and whether or not they could be spillovers, and quantifying their impact on device performance and manufacturing cost; and 3) determining whether these innovations had originated within the LED technology domain, or in a field of research or technology outside of solid-state lighting, making them a technology spillovers.

Relevant  sources for the review were found in an iterative search process that involved two components. The first was the search in specialized patent and publication databases such as Google Scholar, Google Patents, and Scopus , as well as company websites. The second component was the analysis of backward citations in the identified sources, starting from the reviews mentioned in the “Previous literature” section and then iteratively repeating it for all newly identified sources, until no further relevant and significant new sources were found. We also relied on backward citations for the identification of technology spillovers,   considering cited documents as indicators of knowledge origins of an innovation and analyzing whether those documents belonged to the LED technology domain or not.

\subsection{Semi-Structured Interviews}

To supplement our data collection efforts, verify our findings and identify additional spillovers, we conducted a series of elite semi-structured interviews with eleven eminent experts  from academia, industry and public research sector . Experts were selected based on their engagement in  different sub-fields of LED research and manufacturing, as well as the recommendations from other interviewed experts, in essence expanding the list of experts that emerged from the literature review by what is known as a snow-ball sampling method. All interviews were conducted between November 2019 and April 2022 by means of video conferencing and lasted for about one hour. A summary of the background of interviewed experts is provided in Table X in the supplementary information.

The primary, structured part of the interviews explored which innovations were deemed most relevant to the evolution of device performance, consumer experience and manufacturing cost of LED packages. Thereafter, interviewees were asked to consider the extent to which these innovations may have originated outside of their respective field of expertise and the LED industry more broadly—i.e., which of the innovations may be considered spillovers. The remainder of the interview was focused on learning about particular aspects of the manufacturing processes relevant to cost and performance modelling, the current state of industry, and the circumstances surrounding the innovations and spillovers identified in the first part of the interview. Specific quantitative data was also provided by experts, helping fine-tune the parameters of the manufacturing cost model (described in subsection 4.4) and verify device performance data.

### 2023-01-18 2_SSL_EES/article/article_supplementary.tex
added content to main document

>>>>

\section{Anonymized List of Interviewees}

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \# & Sector  & Role  & Country  & Expertise \\ \hline
        1 & R\&D (Academia) & Researcher & UK & Epitaxy \\ \hline
        2 & R\&D (Industry) & Consultant & USA & Device Architecture \\ \hline
        3 & R\&D (Industry) & Head of R\&D & Germany & Epitaxy \\ \hline
        4 & R\&D (Academia) & Professor & Austria & Phosphors \\ \hline
        5 & R\&D (Industry) & Head of R\&D & USA & Device Architecture \\ \hline
        6 & Consulting & Consultant & USA & Device Architecture \\ \hline
        7 & R\&D (Academia) & Professor & Germany & Phosphors \\ \hline
        8 & R\&D (Government) & R\&D Manager  & USA & Device Architecture \\ \hline
        9 & Consulting & Consultant & USA & Device Applications \\ \hline
        10 & R\&D (Academia) & Professor & France & Device Physics \\ \hline
        11 & R\&D (Industry) & Head of R\&D & USA & Device Architecture \\ \hline
        12 & R\&D (Industry) & Head of R\&D & Germany & Phosphors \\ \hline
        13 & R\&D (Industry) & Head of R\&D & USA  & Phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
\end{table}
